---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)

actors <- read_csv("data/actors.csv")
movies <- read_csv("data/movies.csv")
roles <- read_csv("data/roles.csv")

```

```{r}

actors
view(movies)
roles

```
Question 2.

What are the primary keys and the foreign keys for each data frame?
What kind of relationship (one-to-one, one-to-many, many-to-many) is there between
- actors and roles - one to many - pk = ID, fk = ID
- movies and roles - one to many - pk = ID, fk = ID
- actors and movies - many to many - pk = role, fk = actor_id & movie_id

```{r}

# 192017

little_mermaid_roles <- roles %>%
  filter(movie_id == "192017")

```

```{r}

left_join(little_mermaid_roles, actors, by = c("actor_id" = "id"))

```

```{r}

inner_join(little_mermaid_roles, actors, by = c("actor_id" = "id"))

```

```{r}

actors_more_than_once <- actors %>% 
  filter(film_count > 1)

```

```{r}
actors_more_than_once %>%
  left_join(roles, by = c("id" = "actor_id")) %>% 
  left_join(movies, by = c("movie_id" = "id")) %>% 
  mutate(rank = as.numeric(rank)) %>%
  group_by(first_name, last_name) %>% 
  summarise(avg_rank = mean(rank, na.rm = T)) %>%
  ungroup() %>% 
  slice_max(avg_rank, n = 5, with_ties = F)

```


### Busiest actor

```{r}

dataset_complete <- actors_more_than_once %>% 
  left_join(roles, by = c("id" = "actor_id")) %>% 
  left_join(movies, by = c("movie_id" = "id"))


busiest_actor <- dataset_complete %>% 
  group_by(id, first_name, last_name) %>% 
  summarise(num_films = n()) %>% 
  arrange(desc(num_films)) %>% 
  head(1)

```

List the films they appear in, the year the film was released and their role
in the film.

```{r}

semi_join(dataset_complete, busiest_actor, "id")

```
