---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)

single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")

```

```{r}

pattern <- "a"

```

```{r}

single_string %>% 
  str_detect(pattern)

```

```{r}

strings %>% 
  str_detect(pattern)

```

```{r}

pattern <- "str"

strings %>% 
  str_detect(pattern)

```

```{r}

pattern <- "(?i)[A-Z]" #looks for both a and A - ignores case sensitivity.

strings %>% 
  str_detect(pattern)

```

```{r}

# search for number of instances of a character
pattern <- "i{2,}"

strings %>% 
  str_detect(pattern)

```

```{r}

# search "or" patterns
pattern <- "a|d"

strings %>% 
  str_detect(pattern)

```

Make a vector containing the following strings:
"We're having great fun learning about data analysis at CodeClan!"
"CodeClan is at 37 Castle Terrace in Edinburgh and ClockWise in Glasgow"
"Our cohort code starts with DE or DR"
"You get a great view of the castle or the national conservatoire from just outside CodeClan's door"

Write and test regular expressions which will match:
Any string mentioning CodeClan
Any string containing two consecutive capital letters
Any string not including the word "castle" or "Castle"


```{r}

strings <- c("We're having great fun learning about data analysis at CodeClan!",
             "CodeClan is at 37 Castle Terrace in Edinburgh and ClockWise in
             Glasgow",
             "Our cohort code starts with DE or DR",
             "You get a great view of the castle or the national conservatoire
             from just outside CodeClan's door"
)

# Any string mentioning CodeClan
strings %>% 
  str_detect("CodeClan")
  
# Any string containing two consecutive capital letters
strings %>% 
  str_detect("[A-Z]{2}")

# Any string not including the word "castle" or "Castle"
strings %>% 
  str_detect("(?i)castle", negate = TRUE)

```

```{r}

strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)

```


? - zero or one {0,1}
* - zero or more {0,}
+ - one or more {1,}

```{r}

string <- "@emailprovider.com"

pattern <- "[a-z]+@emailprovider.com"

string %>% 
  str_detect(pattern)

```


```{r}

pattern <- "[a-z0-9[:punct:]]+@[a-z[:punct:]]+"
# pattern2 <- "\\.+"  

strings %>% 
  str_extract_all(pattern) %>% 
  flatten_chr()

```

```{r}

pattern <- "[a-z0-9[:punct:]]+@[a-z[:punct:]]+"

strings %>% 
  str_replace_all(pattern, "[REDACTED]")

```
Make a vector containing the strings:
"Colin's mobile number definitely isn't 07777777777."
"A landline might be 01010 123456."
"Mhairi, Del, Aileen and Steph helped write these notes"
"I'm called your_name_here and I live in your_town_here."
Use regular expressions to:
Extract all the names
Extract the landline number
Insert your name and where you live into the fourth string

```{r}
string <- c("Colin's mobile number definitely isn't 07777777777.",
            "A landline might be 01010 123456.",
            "Mhairi, Del, Aileen and Steph helped write these notes",
            "I'm called your_name_here and I live in your_town_here."
            )

# Extract all the names
string %>% 
  str_extract_all("[A-Z][a-z]+") %>% 
  flatten_chr()

# Extract the landline number
string %>% 
  str_extract_all("[0-9]+[:space:][0-9]+") %>%
  flatten_chr()

# Insert your name and where you live into the fourth string
string %>% 
  str_replace("your_name_here", "Chris") %>% 
  str_replace("your_town_here", "Glasgow") 


string %>%
  str_replace(pattern = c("[a-z]+_[a-z]+_[a-z]+","[a-z]+_[a-z]+_[a-z]+"),
                  replacement = c("Jamie", "Glasgow"))

```

```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)
```

```{r}

pattern <- "^[a-z][a-zA-Z]*" # starts with
pattern2 <- "[a-zA-Z]*[a-z]$" # ends with
pattern3 <- "^[a-z]+$" # starts and ends with

test_strings %>% 
  str_extract_all(pattern3)

```

```{r}

student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 

```

```{r}

student_info %>% 
  str_detect(emails, "[a-z0-9]+@")

```

```{r}

student_info %>% 
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))

```

```{r}

student_info %>% 
  mutate(emails = str_extract(emails, "@[a-z.]+"))

```

