---
title: "Session 2 - Missing values"
output: html_notebook
---

```{r}
library(CodeClanData)
library(janitor)
comms_data
```

# Drop all NA's - deletes ALL rows with NA's - lost a lot of data
```{r}
comms_data

comms_data %>%
  drop.na()
```

# Imputation - replace NAs with somthing sensible
maybe the mean, median, range etc
```{r}
comms_data %>%
  pull(monthly_charges) %>% 
  mean(na.rm = T)

comms_data %>%
  pull(monthly_charges) %>% 
  median(na.rm = T)

min_charge <- comms_data %>%
  pull(monthly_charges) %>% 
  min(na.rm = T)

max_charge <- comms_data %>%
  pull(monthly_charges) %>% 
  max(na.rm = T)

min_charge
max_charge

max_charge - min_charge

summary(comms_data)

```

# Leave them alone?

'is.na()'

```{r}
is.na(1)
is.na("1")
is.na(NA)
is.na(NULL)
is.na(NaN)
is.na('unknown')

```

```{r}
comms_data

```

## Standard missing values
'NA'

```{r}
comms_data
comms_data %>% 
  filter(is.na(monthly_charges))

comms_data %>%
  pull(monthly_charges)%>%
  is.na() %>% 
  mean()


comms_data %>% 
  summarise(count = sum(is.na(monthly_charges)),
            mean=mean(is.na(monthly_charges)))

```

```{r}
comms_data %>% 
  drop_na()
```

## Imputation

'coalesce(thing to fix, thing to fix with)'

```{r}
comms_data_imputed_median <- comms_data %>% 
  mutate(monthly_charges)


comms_data %>% 
  select(monthly_charges) %>% 
  mutate(monthly_charges2 = coalesce(monthly_charges,
                                     median(monthly_charges, na.rm = T)))

summary(comms_data)
```

```{r}
comms_data <- clean_names(comms_data)
names(comms_data)

comms_data_imputed_values <- comms_data %>% 
  mutate(monthly_charges2 = coalesce(monthly_charges,
                                     median(monthly_charges, na.rm = TRUE)),
         monthly_charges3 = coalesce(monthly_charges,
                                     mean(monthly_charges, na.rm = TRUE)),
         .after = monthly_charges)

names(comms_data_imputed_values)

```

```{r}

ggplot(comms_data_imputed_values, aes(customer_id, monthly_charges2)) +
  geom_point()

ggplot(comms_data_imputed_values, aes(customer_id, monthly_charges3)) +
  geom_point()
```

# Non-standard missing values

```{r}
comms_data %>% 
  select(payment_method) %>% 
  group_by(payment_method) %>%
  summarise(num_pmt = n()) %>% 
  arrange(desc(num_pmt))

comms_data %>% 
  select(payment_method) %>% 
  group_by(payment_method) %>%
  summarise(num_pmt = n()) %>% 
  arrange(num_pmt)


num_na <- comms_data %>%
  summarise(pmt_na = sum(is.na(payment_method)),
            tc_na = sum(is.na(total_charges)))

num_na
  
```

'na_if()' <-> 'coalesce'

```{r}

comms_data %>% 
  filter(is.na(total_charges))

comms_data

comms_data_imputed_values %>% 
  select(total_charges) %>%  # can delete safely
  mutate(total_charges_explicit_na = na_if(total_charges,"na"),
         total_charges_explicit_na2 = na_if(total_charges_explicit_na,"N/A"))
```

```{r}
library(tidyverse)
library(dplyr)
library(janitor)


comms_data_cleaned <- comms_data %>%
  mutate(total_charges = na_if(total_charges, "na"),
         total_charges = na_if(total_charges, "N/A"),
         payment_method = na_if(payment_method,"--"),
         payment_method = coalesce(payment_method,"Unknown"))

comms_data_cleaned

```

# Recoding

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(older_than_12 = ageyears > 12)

students_big %>% 
  select(ageyears) %>% 
  mutate(older_than_12 = if_else(ageyears >12,
                                 "Older than 12", "12 or younger"))
```

```{r}
students_big %>% 
  select(gender) %>% 
  mutate(gender_recoded = if_else(gender == "M","Male","Female"))
```

```{r}

students_big %>% 
  select(handed) %>% 
  mutate(handed_recoded = if_else(handed == "R", "Right", "Left"))

students_big %>% 
  select(favourite_physical_activity) %>% 
  mutate(fpa_recoded = if_else(favourite_physical_activity == "Football/Soccer",
                               "Football",favourite_physical_activity))

```

```{r}
students_big %>% 
  select(handed, favourite_physical_activity) %>% 
  mutate(handed = if_else(handed == "R", "Right", "Left"),
         favourite_physical_activity = 
           if_else(favourite_physical_activity == "Football/Soccer",
                               "Football",favourite_physical_activity))

```

## `recode()`

```{r}

students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recoded = recode(travel_to_school,
                                           'Bus' = 'bus',
                                           'Car' = 'car',
                                           'Walk' = 'walk'))

students_big  %>% 
  count(travel_to_school)

students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recoded = recode(travel_to_school,
                                           'Bus' = 'bus',
                                           'Car' = 'car',
                                           'Walk' = 'walk',
                                           .default = 'other'))

```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(ageyears_recoded = recode(ageyears, "12" = 120, .default = 0))
```

```{r}

data <- read.table(header = TRUE, text="
 subject sex control cond1 cond2 compl
       1   M     7.9  12.3  10.7  T
       2   F     6.3  10.6  11.1  T
       3   F     9.5  13.1  13.8  N
       4   M    11.5  13.4  12.9  F
")

```

```{r}
 data %>% 
  mutate(subject = recode(subject,
                          "1" = "001",
                          "2" = "002",
                          "3" = "003",
                          "4" = "004"),
         sex = if_else(sex == "M","Male","Female"),
         compl = recode(compl,
                        "T" = TRUE,
                        "F" = FALSE,
                        .default = NA))
```
# NA values must match the value types of the rest of the data set:
i.e. NA_character_ for text based column.


# case_when()

```{r}
x <- 1:10

case_when(x <= 7 ~ "A",
          x <= 5 ~ "B",
          TRUE ~ "C")

```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    age_divisible = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ "Other number"
    )
  )

```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    ageyears = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ as.character(ageyears)
    )
  )
```

# Session 4 - Extra dplyr

```{r}
new_coders

```
# using `group_by()` with `mutate()`

```{r}
new_coders %>% 
  group_by(reasons_to_code) %>% 
  summarise(avg_exp_earn = mean(expected_earn))

new_coders %>% 
  group_by(job_pref) %>% 
  mutate(avg_exp_earn = mean(expected_earn))
```

# Window function

lag() and lead()

```{r}
chinesemeal %>% 
  mutate(last_year_cals = lag(CaloriesPerDay))

chinesemeal %>% 
  mutate(next_year_cals = lead(CaloriesPerDay))

chinesemeal %>% 
  mutate(diff_cals =  CaloriesPerDay - lag(CaloriesPerDay))
```
# `across()`

```{r}
all_deaths



```

```{r}
all_deaths %>% 
  mutate(book1_GoT = as.logical(book1_GoT))

all_deaths %>% 
  summarise(across(book1_GoT:book5_DwD, sum))


```


