---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)

source("../prob.R")

```

```{r}

s_two_dice <- rolldie(2)
s_two_dice

```

```{r}

s_two_dice %>% 
  filter(X1 + X2 == 7)

```

```{r}

s_two_dice <- s_two_dice %>% 
  mutate(total = X1 + X2)

s_two_dice
```

```{r}

s_sum_equals_7 <- s_two_dice %>%
  filter(total == 7)

s_sum_equals_7

```

```{r}


s_sum_gt_10 <- s_two_dice %>%
  filter(total > 10)

s_sum_gt_10

```
# Task

```{r}
# first die even
first_die_even <- s_two_dice %>% 
  filter(X1 %% 2 == 0)

first_die_even
```

```{r}
# die2 minus die 1 is -2
die2_minus_die1 <- s_two_dice %>% 
  filter(X2 - X1 == -2)

die2_minus_die1

```

```{r}
# die values differ by 2
differ_by_two <- s_two_dice %>% 
  filter(abs(X1 - X2) == 2)

differ_by_two
```

# Counting rules

## Single-step experiments

```{r}

s <- rolldie(2)

total_seven <- s %>% 
  filter(X1 + X2 == 7)

total_seven
```

```{r}

nrow(total_seven)/nrow(s)

```

## Multiple-step experiments

Step 1: Rolling a dice - 6 possible outcomes
Step 2: tossing a coin - 2 possible outcomes

Probability of rolling a 6 and tossing a heads:
P(6) = 1/6
P(H) = 1/2


```{r}

s_die <- rolldie(1)
s_coin <- tosscoin(1)

s_total <- merge(s_die, s_coin)

head_six <- s_total %>% 
  filter(X1 == 6 & toss1 == "H")

nrow(head_six) / nrow(s_total)
```

# Permutations and combinations

##Permutation
An arrangement of outcomes in which the _order matters_

##Combination
An arrangement where order _does not matter_

# Task - choose 3 roles for 10 employees 

```{r}

library(gtools)

```

```{r}

nrow(permutations(10, 4, repeats.allowed = FALSE))

```

```{r}

employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Gabra", "Hua", "Inigo", "Hugo")

seniors <- as_tibble(permutations(n = 10, r = 3, v = employees))

seniors <- seniors %>% 
  rename(CEO = V1, CTO = V2, COO = V3)

seniors
```

```{r}

seniors %>% 
  group_by(CEO, CTO) %>% 
  filter(CEO == "Alice" & CTO == "Bob") %>% 
  summarise(probability = n()/nrow(seniors))

```

## Combinations

```{r}

num_draws <- choose(50, 6)
num_draws

1/num_draws
```

# Task

Use the combinations() function in gtools for this task (don’t try to do it manually). I have five flavours of ice cream to choose from:

flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")
and a bowl that can hold three scoops.

How many different variations in bowls of ice cream can I have assuming:

- Each scoop must be a different flavour

```{r}

flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")

comb <- combinations(5, 3, v = flavours, repeats.allowed = F)
num_comb <- nrow(comb)
num_comb

```

- I can repeat flavours if I wish?

```{r}

comb <- combinations(5, 3, v = flavours, repeats.allowed = T)
num_comb <- nrow(comb)
num_comb

```



```{r}
s_four_coins <- tosscoin(4)

s_four_coins <- s_four_coins %>% 
  mutate(num_heads = rowSums(. == "H"))

```

P(2 heads)

```{r}

s_at_least_2_heads <- s_four_coins %>% 
  filter(num_heads >= 2)

#P = 2+ heads
prob_two_heads <- nrow(s_at_least_2_heads) / nrow(s_four_coins)

prob_two_heads
1 - prob_two_heads

```

## "And" combinations

event A: {1,2,3,4}
event B: {3,4,5,6}

```{r}

rolldie(2) %>% 
  filter(X1 %in% 3:4, X2 %in% 3:4)

```


Probability of more than 1 H and 1 T in four coin flips
```{r}

s_four_coins %>% 
  mutate(num_tails = rowSums(. == "T")) %>%  
  filter(num_heads >= 1 & num_tails >= 1) %>% 
  summarise(prob = n()/nrow(s_four_coins))

```

## Mutually exclusive events

```{r}

s_four_coins %>% 
  mutate(num_tails = rowSums(. == "T")) %>%  
  filter(num_heads >= 3 | num_tails >= 2) %>% 
  summarise(prob = n()/nrow(s_four_coins))


```

## OR combination

>union

event A: roll an even number: {2,4,6}
event B: roll one of {4,5}

```{r}

s_two_dice <- rolldie(2)

s_two_dice %>% 
  filter(X1 %% 2 == 0 | X2 == 5) %>% 
  summarise(prob= n()/nrow(s_two_dice))

```

# Task

```{r}

s_two_dice <- rolldie(2)

#first die is even
prob_X1_even <- s_two_dice %>% 
  filter(X1 %% 2 == 0) %>% 
  summarise(prob = n() / nrow(s_two_dice))

#second die is 5
prob_X2_5 <- s_two_dice %>% 
  filter(X2 == 5) %>% 
  summarise(prob = n() / nrow(s_two_dice))

# first die is even AND second die is 5
prob_X1_even_and_X2_5 <- s_two_dice %>% 
  filter(X1 %% 2 == 0 & X2 == 5) %>% 
  summarise(prob = n() / nrow(s_two_dice))

# first die is even OR second die is 5
prob_X1_even_or_X2_5 <- prob_X1_even + prob_X2_5 - prob_X1_even_and_X2_5

prob_X1_even_or_X2_5

```

# Conditional probability

```{r}
library(janitor)
cb <- read_csv("data/colour_blindness.csv")

```

```{r}
cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_totals(where = c("row", "col"))

cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_title()

```

```{r}

prob_female <- cb %>% 
  summarise(prob = sum(bio_sex == "female") / n())

prob_female
```

# Task - marginal probabilities

```{r}
# being male
prob_male <- cb %>% 
  summarise(prob = sum(bio_sex == "male") / n())

prob_male

#colour blind
prob_cb <- cb %>% 
  summarise(prob = sum(cblind) / n())

prob_cb

# not colour blind
prob_not_cb <- cb %>% 
  summarise(prob = sum(cblind == FALSE) / n())

prob_not_cb

```

```{r}

prob_female_cb <- cb %>% 
  filter(bio_sex == "female" & cblind) %>% 
  summarise(prob = n() / nrow(cb))

prob_female_cb

prob_male_cb <- cb %>% 
  filter(bio_sex == "male" & cblind) %>% 
  summarise(prob = n() / nrow(cb))

prob_male_cb
```

