---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)

```





```{r}

date_time1 <- as.POSIXct("2020-01-01")
date_time1

# return seconds of date
unclass(date_time1)

```

```{r}
# list of each time/date component - great for extracting certain values
date_time2 <- as.POSIXlt("2020-01-01")
date_time2
unclass(date_time2)

```

```{r}

netflix <- read_csv("data/netflix.csv")

netflix

```

Column data types:
Date: "<date>"
Time: "<time>"
Date-time: "<dttm>"

```{r}

ymd("2019-01-19")
class(ymd("2019-01-19"))

mdy("January 9th, 2019")
dmy("9-Jan-2019")

ymd_hms("2019-01-19 10:08:26")
ymd_hm("2019-01-19 10:08")
ymd_h("2019-01-19 10")

```

```{r}

netflix <- netflix %>% 
  mutate(date_added = mdy(date_added))

```

## Extracting individual components

```{r}

netflix <- netflix %>% 
  mutate(day = day(date_added),
         month = month(date_added, label = TRUE, abbr = FALSE),
         year = year(date_added),
         weekday = wday(date_added, label = TRUE))

netflix

```

## Calculating time spans

Three types of time span

'Difftime': a datetime object that represents days
'Period': represents a timespan like week, month, day
'Duration': represents an exact number of seconds

```{r}

netflix <- netflix %>% 
  mutate(time_difftime = today() - date_added)

netflix
```

```{r}

netflix <- netflix %>% 
  mutate(time_period = as.period(today() - date_added),
         time_years = as.numeric(time_period, "years"),
         time_hours = as.numeric(time_period, "hours"),
         time_minutes = as.numeric(time_period, "minutes"),
         time_seconds = as.numeric(time_period, "seconds"),
         time_weeks = as.numeric(time_period, "weeks"),
         time_months = as.numeric(time_period, "months"))

netflix

```

### Subtracting periods from dates 

```{r}

netflix <- netflix %>% 
  mutate(promo_date = date_added - weeks(2),
         removal_date = date_added + month(4))

netflix
```
### Durations

```{r}

netflix <- netflix %>% 
  mutate(time_duration = as.duration(today() - date_added))

netflix %>% 
  select(starts_with("time"))

```

can also play around with dseconds(), dminutes(), dhours(), ddays(), dmonths(),
and dyears()

What date is a year after the 10th January 2020

```{r}

ymd(20200110) + years(1)
ymd(20200110) + dyears(1)

```

## Understanding timezones

```{r}
# show all timezones
OlsonNames()

ymd_hms("2023-09-13 10:22:01", tz = "Atlantic/Faeroe")

netflix %>% 
  mutate(date_tz_1 = force_tz(date_added, tzone = "Asia/Singapore"),
         date_tz_2 = with_tz(date_added, tzone = "Asia/Singapore"))

```

# Lab

## Question 1

```{r}

tdf <- read_csv("data/tour_de_france.csv")

tdf
```

## Question 2

```{r}

tdf <- tdf %>% 
  mutate(start_date = mdy(start_date),
         born = dmy(born),
         died = mdy(died))

tdf
```

## Question 3

```{r}

tdf <- tdf %>% 
  mutate(start_month = month(start_date, label = TRUE, abbr = FALSE))

tdf %>% 
  group_by(start_month) %>% 
  summarise(count = n())

```

July is the most popular month to start the TdF.

## Question 4

```{r}

tdf <- tdf %>% 
   mutate(start_wday = wday(start_date, label = TRUE, abbr = FALSE))

tdf %>% 
  group_by(start_wday) %>% 
  summarise(count_wday = n()) %>% 
  slice_max(count_wday)

```
Saturday is the most common week day to start the TdF.

## Question 5

```{r}
# tdf lasts 23 days so event ends on the date 22 days after the start date.
tdf <- tdf %>% 
  mutate(end_date = start_date + days(22))

tdf %>% 
  select(start_date, end_date)
```

## Question 6

```{r}

tdf <- tdf %>% 
  mutate(rider_age_start = as.numeric(as.duration(start_date - born), "years"))

tdf
```

## Question 7

```{r}

depart_time <- ymd_hms("2018-08-07 14:15:00", tz = "Australia/Sydney")

arrive_time <- ymd_hms("2018-08-08 5:40:00", tz = "Europe/London")

depart_time
arrive_time

```

## Question 8 

```{r}

diff_time <- as.numeric(arrive_time - depart_time, "hours")

diff_time

```

## Question 9

```{r}

departure_time <- ymd_hm("2018-08-10 06:20", tz = "Europe/London")
flight_time <- hm("28:46")

arrival_time <- with_tz(departure_time + flight_time, tzone = "Australia/Sydney")

arrival_time
```

```{r}

wee_example <- tibble(
  birthday = ymd("1993-11-19"),
  event_date = ymd("2022-08-22")
)

wee_example %>% 
  mutate(
    age_interval = interval(birthday, event_date),
    age_period = as.period(age_interval, units = "year"),
    age_years = year(age_period)
  )

```

# Introduction to Time Series

```{r}

install.packages(c('tsibble', 'tsibbledata', 'feasts'))
library(tsibble)
library(tsibbledata)

```


"as_tsibble()" converts data to time-series

## "tsibble()"

_index_ : time component
_Key_ : identifies underlying structure of the data, in conjunction with the
_index_ , uniquely identifies each individual row


```{r}

vic_elec

```

```{r}

index(vic_elec)
index_var(vic_elec)

key(vic_elec)
key_vars(vic_elec)

interval(vic_elec)

```

```{r}

vic_elec %>% 
  select(Time, Demand, Temperature)

```

```{r}

vic_elec %>% 
  filter(!Holiday)

vic_elec %>% 
  filter(Time < "2012-01-01 01:00:00")

vic_elec %>% 
  filter(Time < as.POSIXct("2012-01-01 01:00:00", "Australia/Melbourne"))

vic_elec %>% 
  filter(Time < as.Date("2012-01-02"))

vic_elec %>% 
  filter_index("2012-01")

```

```{r}

vic_elec <- vic_elec %>% 
  mutate(year = lubridate::year(Date))

vic_elec
```

```{r}

vic_elec %>% 
  select(year, Temperature) %>% 
  group_by(year) %>% 
  summarise(mean_temp = mean(Temperature))

vic_elec %>% 
  index_by(as_date(Time)) %>% 
  summarise(mean_temp = mean(Temperature))

vic_elec  %>% 
  select(year, Temperature) %>% 
  index_by(year) %>% 
  summarise(mean_temp = mean(Temperature))

```

## "tsibble()" and "ggplot2"

```{r}

vic_elec %>% 
  filter(Holiday == FALSE) %>% 
  ggplot() +
  geom_line(aes(x = Time, 
                y = Temperature))

```

```{r}

floor_date()
ceiling_date()
yearweek()
yearmonth()
yearquarter()

```

```{r}

elec_date <- vic_elec %>% 
  index_by(date = as_date(Time)) %>% 
  summarise(temp_mean = mean(Temperature))

elec_date %>% 
  ggplot(aes(date, temp_mean)) +
  geom_line()


elec_month <- vic_elec %>% 
  index_by(month = month(Time, label = TRUE)) %>% 
  summarise(temp_mean = mean(Temperature))

elec_month %>% 
  ggplot(aes(month, temp_mean)) +
  geom_point() +
  geom_line(group = 1)

elec_year <- vic_elec %>%
  index_by(year = year(Time)) %>%
  summarise(temp_mean = mean(Temperature, na.rm = TRUE))

elec_year %>%
  ggplot(aes(x = year, y = temp_mean)) +
  geom_col(fill = "steelblue", alpha = 0.7) + 
  ylab("Mean Temperature")+ 
  xlab("year")

```

## Rolling windows

```{r}

library(slider)
library(feasts)

```

```{r}

elec_rolling <- vic_elec %>% 
  mutate(
    temp_moving_avg = slide_dbl(
      .x = Temperature,
      .f = ~mean(.x, na.rm = TRUE),
      .before = 1000,
      .after = 1000,
      .complete = TRUE
    ),
    temp_moving_avg_100 = slide_dbl(
      .x = Temperature,
      .f = ~mean(.x, na.rm = TRUE),
      .before = 100,
      .after = 100,
      .complete = TRUE
    )
  )

elec_rolling %>% 
  ggplot() +
  geom_line(aes(x = Date, y = Temperature), col = "gray") +
  geom_line(aes(x = Date, y = temp_moving_avg), col = "red") +
  geom_line(aes(x = Date, y = temp_moving_avg_100), col = "blue") 


```

```{r}

tourism

index_var(tourism)

key_vars(tourism)

tourism %>% 
  distinct(Purpose)

tourism %>% 
  distinct(State)

tourism %>% 
  distinct(Region)

tourism %>% 
  distinct(State, Region, Purpose)

```

```{r}

holidays <- tourism %>% 
  filter(Purpose %in% c("Holiday", "Business")) %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips))

holidays
```

```{r}
holidays %>% 
  autoplot(Trips) +
  xlab("Year")

```

```{r}

holidays %>% 
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>% 
  gg_season(Trips)

```

```{r}

holidays %>% 
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>% 
  gg_subseries(Trips)


```

