---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)

spotify_songs <- read_csv("data/spotify_sample.csv")

```

```{r}

single <- 'this is a single string'
double <- "this is a double string"

single
double

```

```{r}

employees <- tibble(
  first_name = c("Sally", "steven", "Mary", "John"),
  last_name = c("Smith", "hasTings", "Milne", "Davidson"),
  postcode = c("EH10 6GH", "EH3 7HL", "G1 7FR", NA), 
  location = c(" ", "Edinburgh", "", "Stirling")
)

employees

```

```{r}

employees %>%
  group_by(location) %>% 
  summarise(n())

```

```{r}

mary_location <- employees %>% 
  filter(first_name == "Mary") %>% 
  pull(location)



sally_location <- employees %>% 
  filter(first_name == "Sally") %>% 
  pull(location)

```

```{r}

mary_location == sally_location

```

```{r}
employees %>% 
  summarise(nchar(location))

```

```{r}

library(stringr)

```


```{r}

str_c("This","is", "some", "text", sep = " ")
  
```

```{r}

employees %>% 
  mutate(full_name = str_c(first_name, last_name, sep = " "),
         .after = last_name)

```

```{r}
employees %>% 
  mutate(full_name = str_c(first_name, postcode, sep = " "),
         .after = last_name)
```

```{r}
employees %>% 
  mutate(full_name = str_c(first_name, location, sep = " "),
         .after = last_name)
```

```{r}
employees %>% 
  mutate(full_name = paste(first_name, postcode),
         .after = last_name)
```

```{r}

pangram <- "The quick brown fox jumps over the lazy dog"

```

```{r}
# uppercase
pangram %>% 
  str_to_upper

```
```{r}
# lowercase
pangram %>% 
  str_to_lower

```

```{r}
# capitalise first letter of each word in string
pangram %>% 
  str_to_title

```

```{r}
# capitalise first letter of the string
pangram %>% 
  str_to_sentence

```

```{r}

employees %>% 
  select(first_name, last_name) %>% 
  mutate(first_name_fix = str_to_title(first_name),
         last_name_fix = str_to_title(last_name))

```

```{r}

unbroken_string <- "We're going to break up this string"

```

```{r}

unbroken_string %>% 
  str_split(" ") %>% 
  flatten_chr() # remove from list format

```

```{r}

employees %>% 
  mutate(new_postcode = str_split(postcode, " ")) %>%
  unnest(new_postcode)

```

```{r}

employees %>% 
  separate_rows(postcode, sep = " ")

```

```{r}

employees %>% 
  separate(postcode, c("postcode_outer", "postcode_inner"), " ")

```

Concatenate the strings "Manipulating" "strings" "is" "fun!"
Capitalise the result and save it in a new variable
Split the formatted string into individual characters

```{r}

conc_string <- str_c("Manipulating", "strings", "is", "fun!",sep = " ")
upper_string <-  str_to_upper(conc_string)
split_string <- flatten_chr(str_split(upper_string, ""))

conc_string
upper_string
split_string


```

```{r}
full_string <- "Let's grab a substring from this!"

```

```{r}
full_string %>% 
  str_sub(start = 1, end = 5)

```

```{r}
full_string %>% 
  str_sub(start = -5, end = -1)

```

```{r}
employees %>% 
  mutate(postcode_inner = str_sub(postcode, start = -3, end = -1),
         postcode_outer = str_sub(postcode, 1, -4))

```

```{r}

string <- "We're going to replace some words."

```

```{r}

str_sub(string, 1, 22) <- "I've changed"

```

```{r}

string

```

```{r}

number <- 123456.789

```

```{r}

number %>% 
  format(nsmall = 2, decimal.mark = ",")

```
```{r}

number %>% 
  format(big.mark = ",", big.interval = 3)

```

Upcase the text in the playlist_name column. Overwrite the existing column to store the result.
Lowcase the text in the track_artist column. Store the result in a new column called track_artist_lowcase.
Take the first 5 characters of the playlist_id column, and save it in a new column called playlist_start.
Take the first 4 characters of the track_album_release_date column and save it in a new column called year.
What else can you do to this dataset to practice? Can you work with the numbers? This is open ended - have fun!

```{r}
spotify_songs %>% 
  mutate(playlist_name = str_to_upper(playlist_name),
         track_artist_lowcase = str_to_lower(track_artist),
         playlist_start = str_sub(playlist_id, 1, 5),
         year = str_sub(track_album_release_date, 1, 4),
  # additional modifications
         playlist_subgenre = str_to_title(playlist_subgenre),
         tempo = round(tempo, digits = 0),
         duration_ms = format(duration_ms, big.mark = ",", big.interval = 3),
         playlist_genre = str_to_title(playlist_genre))
```

