---
title: "R Notebook"
output: html_notebook
---

```{r}

library(sf)
library(rnaturalearth)
library(leaflet)

```

'{sf}'

stands for simple features

simple means certain geometries are represented by points that don't 
self-intersect (we don't need to worry about this!)

```{r}

north_carolina <- st_read(system.file("shape/nc.shp", package = "sf"))

plot(north_carolina['AREA'])

```

```{r}

nrow(north_carolina)
names(north_carolina)
```

```{r}

nc_geo <- st_geometry(north_carolina)

```

```{r}

nc_geo[[1]]

```

The "big seven" shapes that a feature can be

1. POINT (1 1)
2. MULTIPOINT ((1 1), (2 2))
3. LINE (1 1, 1 2)
4, MULTILINE ((1 1, 1 2), (2 2, 3 3))
5. POLYGON (1 1, 1 2, 2 2, 2 1, 1 1)
6. MULTIPOLYGON ((1 1, 1 2, 2 2, 2 1, 1 1), (1 1, 1 2, 2 2, 2 1, 1 1))
7. GEOMETRY COLLECTION (POINT (1 1), MULTILINE ((1 1, 1 2), (2 2, 3 3)))

Geometry collection can be made up of multiple shape instructions, and any number
of them, just not a geometry collection.

Shetland would be a multipolygon made up of several island polygons.


Can plot individual geometries:

```{r}

plot(nc_geo[[1]])

```

'{sf}' and '{ggplot2}'

with a new geom 'geom_sf'

```{r}
library(tidyverse)

ggplot(north_carolina) +
  geom_sf(aes(fill = AREA), size = 0.1, col = "black") +
  theme_void()

```

```{r}

class(north_carolina)

```

Task 
Try your hand at plotting another feature from the north_carolina dataset. What does it tell you?

```{r}

ggplot(north_carolina) +
  geom_sf(aes(fill = PERIMETER), size = 0.1, col = "black") +
  theme_void()

```

```{r}

names(north_carolina)

```

```{r}

ggplot(north_carolina) +
  geom_sf(fill = "white", colour = "black") +
  geom_sf(data = north_carolina %>%  filter(AREA == max(AREA)), fill = "red")

```

```{r}
north_carolina %>% 
  st_drop_geometry()

```

## Subsetting Simple Features

```{r}

world <- ne_countries(scale = "medium", returnclass = "sf")

names(world)

```


```{r}

world %>% 
  st_drop_geometry() %>% 
  head()

```

```{r}

ggplot(world) +
  geom_sf(aes(fill = pop_est), size = 0.1) +
  scale_fill_viridis_c(trans = "sqrt") +
  theme_void()

```

One thing we can do to "better" visualise the different populations between 
countries is ....> apply a transformation

(India and China are still the most populated countries, but because we reduce how much more populated they are in)

```{r}

pop_estimates <- world$pop_est

```

```{r}

hist(pop_estimates)
hist(sqrt(pop_estimates))
hist(log10(pop_estimates))

```

Task 
Recap your knowledge from ggplot week, and set your geom_sf aesthetic to be filled with the estimated gdp (gdp_md_est variable). Extra points if you make your map colour blind friendly!
What does your plot tell you? What does it tell you compared to the population?

```{r}

world %>% 
  ggplot(aes(fill = gdp_md_est)) +
  geom_sf(size = 0.1) +
  scale_fill_distiller(palette = "Reds", direction = 1, trans = "sqrt") +
  theme_void() +
  labs(fill = "GDP",
       title = "Mid-Year GDP Estimate Heatmap") +
  theme(legend.position = "bottom")


```

Picking out individual countries

```{r}
world %>% 
  filter(name == "United Kingdom") %>% 
  ggplot() +
  geom_sf()
```

Zooming in on particular areas

```{r}
world %>% 
  ggplot(aes(fill = gdp_md_est)) +
  geom_sf(size = 0.1) +
  geom_sf_text(aes(label = name), size = 2, check_overlap = TRUE) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.9)) +
  scale_fill_distiller(palette = "Reds", direction = 1, trans = "sqrt") +
  theme_void() +
  labs(fill = "GDP",
       title = "Mid-Year GDP Estimate Heatmap") +
  theme(legend.position = "bottom")

```

Creating Static Geospatial Data Visualisations

Creating Interactive Geospatial Data Visualisations

'{leaflet}'

```{r}

library(leaflet)

```


Individual Markers

```{r}

leaflet() %>% 
  addTiles() %>% 
  addMarkers(
    lng = 174.768,
    lat = -36.852,
    popup = "Birthplace of R"
  )

```

Circle markers

```{r}

library(jsonlite)

data_url <- "https://data.colorado.gov/resource/j5pc-4t32.json?&county=BOULDER"

water_data <- fromJSON(data_url) %>% 
  jsonlite::flatten(recursive = TRUE)

```

Data preparation

amount, latitude and longitude are not numeric data types, but they should be

```{r}

water_data_clean <- water_data %>% 
  mutate(across(
    c(amount, location.latitude, location.longitude),
    as.numeric
  )) %>% 
  filter(!is.na(location.latitude))

water_data_clean

```

```{r}

leaflet(water_data_clean) %>% 
  addTiles() %>% 
  addCircles(
    lng = ~location.longitude,
    lat = ~location.latitude,
    # make the size depend on the amount of water at that location
    radius = ~amount/10,
    # create data dependent popups - paste or create a string for the labels
    popup = ~paste("water", amount)
  )

```

## Clustering


```{r}


leaflet(water_data_clean) %>% 
  addTiles() %>% 
  addMarkers(
    lng = ~location.longitude,
    lat = ~location.latitude,
    popup = ~paste("water", amount),
    clusterOptions = markerClusterOptions()
  )

```

## Simplifying Shapes
## Reading/Weiting Shapefiles

```{r}

plot(north_carolina['AREA'])

```

```{r}

plot(nc_geo[[4]])

```

```{r}

plot(st_simplify(nc_geo, dTolerance = 2000)[[4]])

```
simplify, increasing the tolerance will decrease the number of points to express
that polygon

by simplifying, we can reduce computational runtime (an important resource in 
shiny apps)

```{r}

north_carolina %>% 
  st_simplify(dTolerance = 2000) %>% 
  ggplot() +
  geom_sf() +
  theme_void()

```

reading/writing shapefiles

```{r}
# not standard to read in a system file, that was mostly for an example
# st_read(system.file())

```

```{r}

north_carolina %>% 
  st_simplify(dTolerance = 2000) %>% 
  st_write(dsn = "nc_simplified",
           layer = "simple_north_carolina",
           driver = "ESRI Shapefile")

```

shapefiles {
- .shp
- .shx
- .prj
- .dbf
}

Usual way we read in shapefiles:

```{r}

north_carolina_simple <- st_read(
  dsn = "nc_simplified/",
  layer = "simple_north_carolina"
  )

```

## Leaflet Polygons

```{r}

pal <- colorBin("Purples", north_carolina_simple$NWBIR74, n = 5)

labels <- paste(
  "<b>", north_carolina_simple$NAME, "</b>", "<br>",
  north_carolina_simple$NWBIR74
) %>%  lapply(htmltools::HTML)

north_carolina_simple %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    weight = 1,
              col = "black",
              fillColor = ~pal(NWBIR74),
              fillOpacity = 100,
              label = labels
    ) %>% 
  addLegend(
    pal = pal,
    values = ~NWBIR74,
    position = "bottomright"
    )

```

## Leaflet in Shiny

