---
title: "R Notebook"
output: html_notebook
---


### Session 2 - Control Flow

# "if" statements

```{r}

if(this_thing_is true) { # THEN
  # do this thing
} else {
  # do this thing instead
}

```

```{r}

x <- 10
y <- 20

if(x < y) {
  print("Yes")
}

```

```{r}

x <- 10
y <- 5:20

if(x %in% y) {
  print("Yes")
}



```

Write an if statement that prints ‘Yes that divides by 2’ if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}

x <- 10

if(x %% 2 == 0){
  print("Yes that divides by 2")
}

```

```{r}

x <- 10

if(x %% 2 == 0){
  print("Yes that divides by 2")
} else {
  print("Nope that does not divide by 2")
}

```

```{r}

x <- 20
y <- 20

if(x %in% y) {
  print("Yes")
} else {
  print("No")
}

```

x = 5, y = 10 -> _"x is in our list" "x is exactly 5"_
x = 6, y = 12 -> _"Start of else statement" "x divides y" "End of else statement"_
x = 6, y = 13 -> _"Start of else statement" "x does not divide Y" "End of else statement"_

```{r}

x <- 5
y <- 10

if (x %in% c(3, 4, 5)){
  
  print("x is in our list")
  
  if (x == 5){
    
    print("x is exactly 5")
    
  }
  
} else {
  
  print("Start of else statement")
  
  if (y %% x == 0){
    
    print("x divides y")
  
  } else{
    
    print("x does not divide Y")
  }
  
  print ("End of else statement")
  
}


```

## Loops


### "for()" loops

- for loops
  - used when we know how many iterations we need
  - will end after desired number of loops

```{r}

v <- 10:20

for (item in v) {
  print(item * 2)
}

```

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (day in weekdays) {
  # browser()   # used to go through loop step by step.
  print(paste("The day is", day))
}

```
### "while()" loops


- while loops
  - used when we don't know how many iterations
  - will end when a condition is met


```{r}

current_value <- 1 # initialise first
stop_value <- 10

while(current_value < stop_value) {
  print(current_value)
  current_value <- current_value + 1
}

# # infinite loop
# while(current_value < stop_value) {
#   print(current_value)
# }

```

### Alternative ways to index for a loop

```{r}

weekdays
len_weekdays <- length(weekdays)

for (i in 1:len_weekdays) {
  phrase <- paste("The day is", weekdays[i])
  print(phrase)
}



```

### Replace a vector from a loop __TERRIBLE PRACTICE__

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
len_weekdays <- length(weekdays)

for (i in 1:len_weekdays) {
  plural <- paste0(weekdays[i], "s") # adding 's' to each vector element
  weekdays[i] <- plural # replacing element from old with new plural element
}

weekdays

```

### Nested for loop

```{r}

m <- matrix(1:12, 4)
nrows <- nrow(m)
ncols <- ncol(m)

for (row_i in 1:nrows) {
  print(m[row_i, ])
}

for (row_i in 1:nrows) {
  for (col_i in 1:ncols) {
      print(m[row_i, col_i])
  }
}

```

```{r}

my_matrix <- matrix(rnorm(25), 5)

num_rows <- dim(my_matrix)[1]
num_cols <- dim(my_matrix)[2]

```

```{r}

product_matrix <- matrix(nrow = 5, ncol = 5)
nrow <- dim(product_matrix)[1]
ncol <- dim(product_matrix)[2]


for (row in 1:nrow) {
  for(col in 1:ncol){
    product_matrix[row, col] <- row * col
  }
}

product_matrix

```

```{r}

example_matrix = matrix(1:25, nrow = 5, ncol = 5)
example_matrix 

for (row_index in 1:5){
  for (col_index in 1:5){
    if (row_index == col_index){
      print(example_matrix[row_index, col_index])
    }
   }
}

```

