---
title: "R Notebook"
output: html_notebook
---

# Session 1 - Matrices and Lists


```{r}

library(tidyverse)

```


```{r}

x <- list(1, "a", TRUE, c(1,2,3))

# x

v <- 1:5

# v

c(1, "a", TRUE, c(1,2,3))

# v

as.list(v)

```

```{r}

x[[4]]
x[4]
v[[2]]
v[2]

glimpse(x)
str(x)

```

### Named List

```{r}

named_list <- list(a = c("I love R","hi"), b = 1:10)

named_list

str(named_list)

named_list$a
named_list$b

names(named_list)

# using list index
named_list[[1]] # equivalent to pull function in tidyverse.
named_list[1]   # equivalent to select function in tidyverse.

# using list name identifier
named_list[['a']]
named_list['a']


```

```{r}

student_info <- list(
                      "Agnes" = list(height = 1.8, shoe_size = 8L),
                      "Bobby" = list(height = 2.1, gender = "Male"),
                      "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info
str(student_info)

student_info$Agnes$height
student_info[["Agnes"]][["height"]]

n <- "Agnes"
student_info$Agnes$height       #use this if you are sure which column you want and it won't change
student_info[[n]][["height"]]   # use this if you want to access programmatically, it's also safer if list names change


```

    Make a named list that contains a numeric vector and a character vector.
    Access the second element of your list using double square brackets [[]]. What is it’s class?
    Access the second element of your list using single square brackets []. What is it’s class?
    Access the second element of your list by name. What is it’s class?

```{r}

v <- list(num = 1, chr = "a")

v[[2]]
class(v[[2]])
# outputs class is character

v[2]
class(v[2])
# output class is list

class(v[["chr"]])
class(v$chr)
# output class is character

```

### Matrices

dataframes are _heterogeneous_ data structures
vectors/matrices are _homogeneous_ data structures

### Creating Matrices

```{r}

m <- matrix(1:12, ncol = 3, byrow = F)
v <- 1:5
v

class(m)
typeof(m)


```

# Indexing a matrix

```{r}

m[1, ] # select first row
m[, 1] # select first col

class(m[1, ])
is.vector(m[1, ])

m[1,2] # select individual value
m[2:3, ] # returns rows 2 and 3 and every column
m[c(1,3), c(1,3)] # select specific rows and columns
m[c(1,3), -2] # select specific rows and columns (minus sign removes 2nd col)

v <- 10:15
v[-2]

```

```{r}

v * 2

m * 2

```

```{r}

m
rowSums(m)
colSums(m)

rowMeans(m)
colMeans(m)

```

Task - 5 minutes
    Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
    Find the element in the 2nd row, 3rd column.
    Check that all the columns are identical by checking that their sums are the same.
    
```{r}

m <- matrix(1:3, nrow = 3, ncol = 4)
m

m[2,3]
colSums(m)

colnames(m) <- paste0("col_", 1:4)
rownames(m) <- paste0("row_", 1:3)
m["row_3", "col_2"]

```
